Ultrafast Electron Microscopes: Design Criteria, Electron Sources, and Column Modeling

### 2013-04-12 inc/hardware/laser/laser.tex
massive spellchecking

The Watt-level green output power from the frequency-doubled sub-picosecond Yb:KGW oscillator corresponds to a visible peak laser pulse power in excess of 100kW, which is well-suited for further frequency conversion through harmonic generation or pumping of an ultrashort pulse optical parametric oscillator.
To our knowledge, the generation of subpicosecond green peak pulse powers greater than 100kW by frequency doubling the output of a laser oscillator has only been demonstrated for high-power thin-disk~\cite{Brunner_powerful_2004,Marchese_pulse_energy_06} and cavity-dumped~\cite{Palmer_microjoule_07} Yb-doped solid-state lasers, although externally-doubled commercial Ytterbium femtosecond oscillators~\cite{website_amplitude,website_solar,website_high_q,website_time_bandwidth} and oscillators with rod-like Yb-fiber gain media~\cite{Ortac_pulse_2007} can be expected to yield comparable peak green pulse powers.

@@

\begin{figure}
  \centering
  \includegraphics{performance.jpg}
  \caption[Power measuments of employed ultrast laser]{
Performance of the diode-pumped, TLS, Yb:KGW laser (in the cavity configuration of \protect\ref{fig:laser-cavity}\protect\subref{fig:laser_schematic} ) as a function of applied diode current: continuous-wave (CW) output power (black squares); mode locked (ML) output power (red circles); frequency doubled power (green triangles); and second harmonic generation (SHG) efficiency (line).
  }
  \label{fig:laser-performance}
\end{figure}

@@

The ML output from the femtosecond Yb:KGW laser with a net negative GVD of $-5,200\text{fs}^2$ per round-trip was intially frequency doubled in a critically phase-matched, 2mm, Type I Lithium triborate (LBO) crystal.
The crystal is uncoated and cut for frequency doubling with the fundamental $p$-polarized radiation at 1046nm incident at Brewster's angle --- a geometry that ensures both minimum loss for the fundamental radiation and, more importantly, a high damage threshold.
The doubling crystal is positioned at the focus of a $z$-fold comprising a 1040nm high reflector with a radius of curvature of 75mm followed by a collimating mirror with a radius of curvature of 100mm reflective at both the fundamental and second harmonic wavelengths.
At the 3.5W ML output power, with a pulse duration of 250fs, the estimated focal irradiance for the fundamental is $\sim50\text{GW}/\text{cm}^2$ in the LBO doubling crystal.
Two plane dispersion compensation mirrors, each with a GVD of $-220\text{fs}^2$ (Layertec GmbH~\cite{website_layertec}), inserted before the $z$-fold doubling geometry compensate for propagation through the 6mm-thick fused silica output coupler, reflections off standard optics, etc.
After the collimating
%page 8639
mirror, a standard flat dichroic beamsplitter then separates the reflected $s$-polarized 523nm radiation from the transmitted $p$-polarized fundamental.

>>>>

The Watt-level green output power from the frequency-doubled sub-picosecond Yb:KGW oscillator corresponds to a visible peak laser pulse power in excess of 100kW, which is well-suited for further frequency conversion through harmonic generation or pumping of an ultrashort pulse optical parametric oscillator.
To our knowledge, the generation of sub-picosecond green peak pulse powers greater than 100kW by frequency doubling the output of a laser oscillator has only been demonstrated for high-power thin-disk~\cite{Brunner_powerful_2004,Marchese_pulse_energy_06} and cavity-dumped~\cite{Palmer_microjoule_07} Yb-doped solid-state lasers, although externally-doubled commercial Ytterbium femtosecond oscillators~\cite{website_amplitude,website_solar,website_high_q,website_time_bandwidth} and oscillators with rod-like Yb-fiber gain media~\cite{Ortac_pulse_2007} can be expected to yield comparable peak green pulse powers.

@@

\begin{figure}
  \centering
  \includegraphics{performance.jpg}
  \caption[Power measurements of employed ultrafast laser]{
Performance of the diode-pumped, TLS, Yb:KGW laser (in the cavity configuration of \protect\ref{fig:laser-cavity}\protect\subref{fig:laser_schematic} ) as a function of applied diode current: continuous-wave (CW) output power (black squares); mode locked (ML) output power (red circles); frequency doubled power (green triangles); and second harmonic generation (SHG) efficiency (line).
  }
  \label{fig:laser-performance}
\end{figure}

@@

The ML output from the femtosecond Yb:KGW laser with a net negative GVD of $-5,200\text{fs}^2$ per round-trip was initially frequency doubled in a critically phase-matched, 2mm, Type I Lithium triborate (LBO) crystal.
The crystal is uncoated and cut for frequency doubling with the fundamental $p$-polarized radiation at 1046nm incident at Brewster's angle --- a geometry that ensures both minimum loss for the fundamental radiation and, more importantly, a high damage threshold.
The doubling crystal is positioned at the focus of a $z$-fold comprising a 1040nm high reflector with a radius of curvature of 75mm followed by a collimating mirror with a radius of curvature of 100mm reflective at both the fundamental and second harmonic wavelengths.
At the 3.5W ML output power, with a pulse duration of 250fs, the estimated focal irradiance for the fundamental is $\sim50\text{GW}/\text{cm}^2$ in the LBO doubling crystal.
Two plane dispersion compensation mirrors, each with a GVD of $-220\text{fs}^2$ (Layertec GmbH~\cite{website_layertec}), inserted before the $z$-fold doubling geometry compensate for propagation through the 6mm-thick fused silica output coupler, reflections off standard optics, etc.
After the collimating
%page 8639
mirror, a standard flat dichroic beamsplitter then separates the reflected $s$-polarized 523nm radiation from the transmitted $p$-polarized fundamental.

### 2013-04-12 inc/hardware/magnetic_lenses/maglens.tex
massive spellchecking

Discussion of the formal mathematics of magnetic lenses can be found in Section \ref{sec:mag_lens_model}.
While in principle one could design a magnetic lens using \ref{eq:lens_eq_of_motion} and \ref{eq:field_of_loop}, the process wouldn't be very practical or instructive.
Further --- as suggested by the use of the auxiliary field $H$ field in \ref{eq:field_of_loop} --- better performance of the lens can be achieved by concentrating the field using materials with high magnetic permeablility.
These materials, called ``magnetic shielding'' since they capture the magnetic field lines, can direct the field lines to a small gap, known as the ``pole-piece gap''.
The field lines exit the shielding in the gap and it is this field that acts as the lens.
This gap can be much smaller than the total width of the coil turns, and yet contain a field strength commensurate with that of the full coil.
This geometry is presented schematically as \ref{fig:polepiece-schematic}.

@@

The high magnetic permeability shielding is made of eleven pieces.
There are two end pieces which are revolution solids of an ``L'' shaped profile.
These ends slide together over the drift tube until touching the disk at the midpoint; the disk is of the proper outer diameter to match the sheilding's smaller outer diameter.
The combination of the two ends and this disk make a ``U'' shape which contains the windings.
The top of the each end's large outer disk (the walls of the winding container) are flattend on eight sides.
These are the mount points for eight bridge pieces which connect the ends to complete the field-capturing shielding loop.
The gaps between the shielding bridges also allow for direct air cooling of the magnetic coils (see Section \ref{sec:mag_lens_thermal}).

@@

%TODO need reference for equation and \alpha_{Cu}
As the lenses have no active cooling system, it was prudent to characterize the thermal response to a moderate current load.
Resistance of metals changes as a function of temperature by $ R = R_0 [ 1 + \alpha ( T - T_0 ) ] $, where $\alpha$ is a material contant; for copper, $\alpha = 3.9 * 10^{-3} \text{K}^{-1}$. 
One can calculate the temperature increase of the system by monitoring the change in resistance.
Maintaining a constant current of 4A through the lens, the voltage drop increased from 1.22V ($R_0 = 0.305\Omega$) to 1.25V ($R=0.3125\Omega$) as the lens warmed.
The resultant temperature increase is only $T - T_0 = 2\text{K}$.
This result is easily confirmed manually; the oxygen-free copper wire coils remain cool to the touch under these operating conditions.

>>>>

Discussion of the formal mathematics of magnetic lenses can be found in Section \ref{sec:mag_lens_model}.
While in principle one could design a magnetic lens using \ref{eq:lens_eq_of_motion} and \ref{eq:field_of_loop}, the process wouldn't be very practical or instructive.
Further --- as suggested by the use of the auxiliary field $H$ field in \ref{eq:field_of_loop} --- better performance of the lens can be achieved by concentrating the field using materials with high magnetic permeability.
These materials, called ``magnetic shielding'' since they capture the magnetic field lines, can direct the field lines to a small gap, known as the ``pole-piece gap''.
The field lines exit the shielding in the gap and it is this field that acts as the lens.
This gap can be much smaller than the total width of the coil turns, and yet contain a field strength commensurate with that of the full coil.
This geometry is presented schematically as \ref{fig:polepiece-schematic}.

@@

The high magnetic permeability shielding is made of eleven pieces.
There are two end pieces which are revolution solids of an ``L'' shaped profile.
These ends slide together over the drift tube until touching the disk at the midpoint; the disk is of the proper outer diameter to match the shielding's smaller outer diameter.
The combination of the two ends and this disk make a ``U'' shape which contains the windings.
The top of the each end's large outer disk (the walls of the winding container) are flattened on eight sides.
These are the mount points for eight bridge pieces which connect the ends to complete the field-capturing shielding loop.
The gaps between the shielding bridges also allow for direct air cooling of the magnetic coils (see Section \ref{sec:mag_lens_thermal}).

@@

%TODO need reference for equation and \alpha_{Cu}
As the lenses have no active cooling system, it was prudent to characterize the thermal response to a moderate current load.
Resistance of metals changes as a function of temperature by $ R = R_0 [ 1 + \alpha ( T - T_0 ) ] $, where $\alpha$ is a material constant; for copper, $\alpha = 3.9 * 10^{-3} \text{K}^{-1}$. 
One can calculate the temperature increase of the system by monitoring the change in resistance.
Maintaining a constant current of 4A through the lens, the voltage drop increased from 1.22V ($R_0 = 0.305\Omega$) to 1.25V ($R=0.3125\Omega$) as the lens warmed.
The resultant temperature increase is only $T - T_0 = 2\text{K}$.
This result is easily confirmed manually; the oxygen-free copper wire coils remain cool to the touch under these operating conditions.

### 2013-04-12 inc/model_results/lens_charge/lens_charge.tex
massive spellchecking

In the transverse spatial dimension, space-charge effects that increase $ \sigma_{ \smallT } $ can be compensated for by using magnetic electron lenses~\cite{oudheusden_electron_2007,lagrange_nanosecond_2008}.
In an UEM, of course, the electron pulse will also need to be focused onto the specimen --- the focal spot size, post specimen image magnification, and CCD detector pixelation determining the spatial resolution in a TEM geometry~\cite{berger_dc_2009}.
The extent to which space-charge effects limit the focal fidelity is expected to be dependent upon both $N$ and the pulse shape (i.e., the pulse's space-time eccentricity), while still subject to Liouville's Theorem (see Section \ref{sec:liouville}).

@@

In the preceding sections it has already been established that oblate eccentricities are preferred for UEM in general.
When focusing the pulse it can be shown that oblate pulses will also have markedly better performance.
\ref{fig:focus_lens_charge} shows the predicted dynamics of the spatial electron pulse waist (HW1/eM) for intially oblate ($ \xi ( 0 ) = 0.1 $) or prolate ($ \xi ( 0 ) = 10 $) pulses under perfect focusing for lenses of two focal lengths; a `long' focal length of 60mm and a `short' focal length of 6.0mm.
The pulse's initial volume is constant in all cases ($w_{\smallT}^2 w_{z} = (232 \mu\text{m})^3$).
For both lens strengths, the deleterious effects of space-charge are evident as the number of electrons/pulse $ N $ is increased.
To make an effective comparison, highlighting the effect of space-charge alone, this simulation again uses the idealistic initial conditions set; the pulses used are at velocity associated with a 30kV energy and have momentum variance associated with $\Delta E = $ 0.5eV, but are not simulated using an emitter and accelerator.
As such the simulation presented approximates a beam which has been nearly perfectly collimated at the focus of a long focal-length lens.

>>>>

In the transverse spatial dimension, space-charge effects that increase $ \sigma_{ \smallT } $ can be compensated for by using magnetic electron lenses~\cite{oudheusden_electron_2007,lagrange_nanosecond_2008}.
In an UEM, of course, the electron pulse will also need to be focused onto the specimen --- the focal spot size, post specimen image magnification, and CCD detector pix elation determining the spatial resolution in a TEM geometry~\cite{berger_dc_2009}.
The extent to which space-charge effects limit the focal fidelity is expected to be dependent upon both $N$ and the pulse shape (i.e., the pulse's space-time eccentricity), while still subject to Liouville's Theorem (see Section \ref{sec:liouville}).

@@

In the preceding sections it has already been established that oblate eccentricities are preferred for UEM in general.
When focusing the pulse it can be shown that oblate pulses will also have markedly better performance.
\ref{fig:focus_lens_charge} shows the predicted dynamics of the spatial electron pulse waist (HW1/eM) for initially oblate ($ \xi ( 0 ) = 0.1 $) or prolate ($ \xi ( 0 ) = 10 $) pulses under perfect focusing for lenses of two focal lengths; a `long' focal length of 60mm and a `short' focal length of 6.0mm.
The pulse's initial volume is constant in all cases ($w_{\smallT}^2 w_{z} = (232 \mu\text{m})^3$).
For both lens strengths, the deleterious effects of space-charge are evident as the number of electrons/pulse $ N $ is increased.
To make an effective comparison, highlighting the effect of space-charge alone, this simulation again uses the idealistic initial conditions set; the pulses used are at velocity associated with a 30kV energy and have momentum variance associated with $\Delta E = $ 0.5eV, but are not simulated using an emitter and accelerator.
As such the simulation presented approximates a beam which has been nearly perfectly collimated at the focus of a long focal-length lens.

### 2013-04-12 inc/model_results/lens_charge/lens_charge_plot.tex
improve lens_charge multiplot
killing TODOs
more corrections and punctuation
massive spellchecking

%TODO fix axes to be more consistent
%TODO add N=1e8 for oblate?

\begin{figure}
  \centering
  \centerline{
    \subfloat[][] {
      \label{fig:focus_long_oblate}
      \begin{tikzpicture}
        \input{lens_60mm_oblate_data}
        \node at (5.9,1.65) {$N=10^5$};
        \node at (4,3) {$N=10^6$};
        \node at (3,4) {$N=10^7$};
      \end{tikzpicture}
    }
    \subfloat[][] {
      \label{fig:focus_long_prolate}
      \begin{tikzpicture}
        \input{lens_60mm_prolate_data}
        \node at (4.2,2.3) {$N=10^5$};
        \node at (3.3,3.5) {$N=10^6$};
      \end{tikzpicture}
    }
  }
  \centerline{
    \subfloat[][] { 
      \label{fig:focus_short_oblate}
      \begin{tikzpicture}
        \input{lens_6mm_oblate_data}
        \node at (6.5,3) {$N=10^7$};
      \end{tikzpicture}
    }
    \subfloat[][] { 
      \label{fig:focus_short_prolate}
      \begin{tikzpicture}
        \input{lens_6mm_prolate_data}
        \node at (6.9,2.6) {$N=10^7$};
        \node at (6,2) {$N=10^6$};
      \end{tikzpicture}
    }
  }
  \caption[AG model simulated focal behavior of perfect magnetic lenses as a function of pulse charge $N$]{
    AG model simulated focal behavior of perfect magnetic lenses as a function of pulse charge $N$.
    Each plot's pulses begin at the lens' front focal plane at 20 kV velocity ($c/3$) and have the same volume.
    The left plots (\subref{fig:focus_long_oblate} and \subref{fig:focus_short_oblate}) are initially oblate shaped, HW1/eM width $ 500 \mu \text{m}$ and ellipticiy $ \xi ( 0 ) = 0.1 $.
    The right plots (\subref{fig:focus_long_prolate} and \subref{fig:focus_short_prolate}) are initially prolate shaped, HW1/eM width $ 107.7 \mu \text{m}$ and ellipticiy $ \xi ( 0 ) = 10 $.
    Distance traveled in column is measured relative to the lens at $z=0$, in units of the focal lenth.
    The top plots (\subref{fig:focus_long_oblate} and \subref{fig:focus_long_prolate}) have a longer focal length $f = $ 60mm compared to the bottom plots (\subref{fig:focus_short_oblate} and \subref{fig:focus_short_prolate}) $ f = $ 6.0mm.
    A logarithmic scale is used for comparative clarity near the foci.
    Clearly the best performance is achieved for shorter focal lengths and oblate pulses \subref{fig:focus_short_oblate} even at higher initial charge densities.
  }
  \label{fig:focus_lens_charge}
\end{figure}

>>>>

\begin{figure}
  \centering
  \centerline{
    \subfloat[][] {
      \label{fig:focus_long_oblate}
      \begin{tikzpicture}
        \input{lens_60mm_oblate_data}
        \node at (2.7,4.9) [align=left] {$\xi=0.1$\\$f=$ 60mm};
        \node at (5.9,1.65) {$N=10^5$};
        \node at (4.5,2.7) {$N=10^6$};
        \node at (4,3.9) {$N=10^7$};
      \end{tikzpicture}
    }
    \subfloat[][] {
      \label{fig:focus_long_prolate}
      \begin{tikzpicture}
        \input{lens_60mm_prolate_data}
        \node at (2.7,4.9) [align=left] {$\xi=10$\\$f=$ 60mm};
        \node at (4.2,2.3) {$N=10^5$};
        \node at (3.3,3.5) {$N=10^6$};
      \end{tikzpicture}
    }
  }
  \centerline{
    \subfloat[][] { 
      \label{fig:focus_short_oblate}
      \begin{tikzpicture}
        \input{lens_6mm_oblate_data}
        \node at (2.7,4.9) [align=left] {$\xi=0.1$\\$f=$ 6mm};
        \node at (6.5,3) {$N=10^7$};
      \end{tikzpicture}
    }
    \subfloat[][] { 
      \label{fig:focus_short_prolate}
      \begin{tikzpicture}
        \input{lens_6mm_prolate_data}
        \node at (2.7,4.9) [align=left] {$\xi=10$\\$f=$ 6mm};
        \node at (6.9,2.6) {$N=10^7$};
        \node at (6,2) {$N=10^6$};
      \end{tikzpicture}
    }
  }
  \caption[AG model simulated focal behavior of perfect magnetic lenses as a function of pulse charge $N$]{
    AG model simulated focal behavior of perfect magnetic lenses as a function of pulse charge $N$.
    Each plot's pulses begin at the lens' front focal plane at velocity $c/3$ (related to 30 kV energy) and have the same volume.
    The left plots (\subref{fig:focus_long_oblate} and \subref{fig:focus_short_oblate}) are initially oblate shaped, HW1/eM width $ 500 \mu \text{m}$ and ellipticity $ \xi ( 0 ) = 0.1 $.
    The right plots (\subref{fig:focus_long_prolate} and \subref{fig:focus_short_prolate}) are initially prolate shaped, HW1/eM width $ 107.7 \mu \text{m}$ and ellipticity $ \xi ( 0 ) = 10 $.
    Distance traveled in column is measured relative to the lens at $z=0$, in units of the focal length.
    The top plots (\subref{fig:focus_long_oblate} and \subref{fig:focus_long_prolate}) have a longer focal length $f = $ 60mm compared to the bottom plots (\subref{fig:focus_short_oblate} and \subref{fig:focus_short_prolate}) $ f = $ 6.0mm.
    A logarithmic scale is used for comparative clarity near the foci.
    Clearly the best performance is achieved for shorter focal lengths and oblate pulses \subref{fig:focus_short_oblate} even at higher initial charge densities.
  }
  \label{fig:focus_lens_charge}
\end{figure}

### 2013-04-12 inc/model_results/spacecharge/spacecharge.tex
killing TODOs
massive spellchecking

As seen in Section \ref{sec:initial_shapes}, at low pulse charge densities, the propagation dynamics should be largely determined by the effects associated with the intrinsic geometric pulse broadening; for most photoemission processes this is related to the excess photoemission energy $\Delta E$.
On the other hand, space-charge effects are expected to dominate at large pulse charge densities.
A full analysis of how high the charge density may be before deleterious space-charge become unmanagable would be a sizable task.
While much discussion could be had on the degree of distortion allowable, as well as considerations of the initial shape of the pulse and length, and arrangement of the column and its constituent elements, a certain amount of intuition can be gained simply by considering a progressive series of examples.

\begin{figure}
  \centering
  \begin{tikzpicture}
    \input{spacecharge_data_noacc}
  \end{tikzpicture}
  \caption[Idealistic free-space pulse evolution vs charge density]{
    The final HW1/eM pulse width (solid) and length (dashed) after a 15 cm propagation. 
    Pulses are intially similar except for the number of electrons in the pulse, each having 1 mm (HW1/eM) width and 1 ps in duration travelling at $c/3$ with 0.5eV excess photoemission energy.
  }
  \label{fig:spacecharge_noacc}
\end{figure}

@@

%TODO highlight the following in the experimental section
%In other words, the spatial and temporal dispersive effects due to the nonzero value of $\Delta E$ are expected to dominate the space-time dynamics of the electron pulse during its propagation
%in our investigations because less than 10 5 electrons/pulse
%are generated.

@@

\begin{figure}
  \centering
  \begin{tikzpicture}
    \input{spacecharge_data_acc}
  \end{tikzpicture}
  \caption[More realistic free-space pulse evolution vs charge density]{
    The final HW1/eM pulse width (solid) and length (dashed) after a 15 cm propagation. 
    Pulses are intially similar except for the number of electrons in the pulse, each is generated by 1 mm (HW1/eM) width 1 ps laser pulses with 0.5eV excess photoemission energy and then traverses a 20mm acceleration region followed by drift region, accelerating to a velocity of $c/3$.
  }
  \label{fig:spacecharge_acc}
\end{figure}

>>>>

As seen in Section \ref{sec:initial_shapes}, at low pulse charge densities, the propagation dynamics should be largely determined by the effects associated with the intrinsic geometric pulse broadening; for most photoemission processes this is related to the excess photoemission energy $\Delta E$.
On the other hand, space-charge effects are expected to dominate at large pulse charge densities.
A full analysis of how high the charge density may be before deleterious space-charge become unmanageable would be a sizable task.
While much discussion could be had on the degree of distortion allowable, as well as considerations of the initial shape of the pulse and length, and arrangement of the column and its constituent elements, a certain amount of intuition can be gained simply by considering a progressive series of examples.

\begin{figure}
  \centering
  \begin{tikzpicture}
    \input{spacecharge_data_noacc}
  \end{tikzpicture}
  \caption[Idealistic free-space pulse evolution vs charge density]{
    The final HW1/eM pulse width (solid) and length (dashed) after a 15 cm propagation. 
    Pulses are initially similar except for the number of electrons in the pulse, each having 1 mm (HW1/eM) width and 1 ps in duration traveling at $c/3$ with 0.5eV excess photoemission energy.
  }
  \label{fig:spacecharge_noacc}
\end{figure}

@@

@@

\begin{figure}
  \centering
  \begin{tikzpicture}
    \input{spacecharge_data_acc}
  \end{tikzpicture}
  \caption[More realistic free-space pulse evolution vs charge density]{
    The final HW1/eM pulse width (solid) and length (dashed) after a 15 cm propagation. 
    Pulses are initially similar except for the number of electrons in the pulse, each is generated by 1 mm (HW1/eM) width 1 ps laser pulses with 0.5eV excess photoemission energy and then traverses a 20mm acceleration region followed by drift region, accelerating to a velocity of $c/3$.
  }
  \label{fig:spacecharge_acc}
\end{figure}
